%!TEX root = ../dissertation.tex
% the acknowledgments section
My parents built the foundation which I walk upon. My mother gave me strength whenever I needed. My father never spared efforts with respect to my education.

I would like to thank my thesis supervisor, Dr. Marcelo Reyes, for his invaluable orientation. In special for his always far-sighted planning, that alleviated my restlessness and propelled the project to closures. I would also like to thank Gabriela Chiuffa for the collection of experimental data and discussions, and Eliezyer Fermino for his own collection of experimental data. The project would unquestionably falter without their most technical work with the subjects. It would also be impossible without subjects: I want to acknowledge the rats that made part in the experiments.

I would like to give special thanks to Gabriela Melo, for her utmost support. For reading through and commenting lots of drafts, helping edit and think through, and for inspiring me when conclusions seemed afar.

I am very grateful of the companionship of and discussions with my colleagues. Fernanda, Esaú, Pedro, Vanessa, Attalya, Louise, Janine, Estela, Mateus Nogueira, Mateus Silvestrin, Camila, Brenda, Carlos. It is hard to pinpoint the exact contributions, but not as hard as imagining the process without their everyday comments. Some professors were specially helpful in this journey, giving input about the scientific process, about specific questions regarding theoretical challenges, or even about the academy and its bureaucracies. André Cravo, Claudinei Biazoli, Yossi Zana, Rodrigo Pavão and Marcelo Caetano, thank you very much.

I want to thank my work supervisor, Dr. Renato Vicente, for believing in me, sparking creativity, and challenging my limitations. He helped me step up over the fog around to see beyond.

The project received financial support of the Coordination for the Improvement of Higher Education Personnel (CAPES), under financing code 001. It also received support of the Serasa Experian DataLab with time committed to research.

% \textcolor{red}{This is the dedication. Incluir o texto que menciona o apoio da Coordenação de Aperfeiçoamento de Pessoal de Nível Superior - Brasil (CAPES) - Código de Financiamento 001 no trabalho do discente, conforme Portaria CAPES nº 206 de 04/09/2018.}